---
title: "Red mullet fishery data"
author: "moi"
date: "11/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library
require(dplyr)
require(ggplot2)
require(kableExtra)
require(sf)
require(tidyr)
```

# Framework

A short report about fishery data and red mullet.

# Landings

Landings from the [SACROIS system](https://sih.ifremer.fr/Debarquements-effort-de-peche/Sacrois).

```{r saclan}
#require(arrow)
#ds<-open_dataset("/home/moi2/ifremer/dataCOST/parquet/simplesacrois")
#french, fao code for red mullet: MUR, area 27
#navday<-ds%>%filter(PAVILLON=="FRA",ESP_COD_FAO=="MUR",grepl("27.",SECT_COD_SACROIS_NIV3))%>%collect()
#saveRDS(navday,file="../data_confidential/navday.rds")
#load the saved data
#navday<-readRDS("../data_confidential/navday.rds")
```

## Exploratory analysis

### Data summary

Computation of landings (in tons) and effort (days at sea) by areas, rectangle and fishing gear.

```{r datasummary}
#classloa<-c("u10","10-12","12-15","15-18","18-24","24-40","o40")
#lan<-navday%>%
#	group_by(SEQ_ID,
#		 MAREE_ID,
#		 area=SECT_COD_SACROIS_NIV3,
#		 rect=SECT_COD_SACROIS_NIV5,
#		 year=AN,
#		 gear=substr(METIER_DCF_6_COD,1,3),
#		 classloa=classloa[1+findInterval(LONGUEUR_HT,vec=c(10,12,15,18,24,40))]
#		 )%>%
#	summarise(lan_tot=sum(QUANT_POIDS_VIF_SACROIS,na.rm=T)/1000)%>%
#	ungroup()
#eff<-navday%>%
#	select(SEQ_ID,MAREE_ID,fday=totFishDays_seqs)%>%
#	distinct()%>%
#	group_by(SEQ_ID,MAREE_ID)%>%
#	summarise(fday=sum(fday,na.rm=T))%>%
#	ungroup()
#laneff<-full_join(lan,eff)%>%group_by(year,area,rect,gear,classloa)%>%
#	summarise(lan_tot=sum(lan_tot),fday=sum(fday))%>%
#	ungroup()
#saveRDS(laneff,file="../data_confidential/laneff.rds")
laneff<-readRDS(file="../data_confidential/laneff.rds")

```


### Landings by area

Landings in ton & percentage, with fishing effort over the period 2000-2022 by area.

```{r tab1}
tab1<-laneff%>%group_by(area)%>%
	summarise(lan_tot=sum(lan_tot),fday=sum(fday))%>%
	mutate(perc_lan=round(100*lan_tot/sum(lan_tot),2))%>%
	ungroup()%>%
	arrange(desc(perc_lan))
kable(tab1)
```

Landings in ton & percentage, with effort over the period 2000-2022 by fishing gear. 

```{r tab2}
tab2<-laneff%>%group_by(gear)%>%
	summarise(lan_tot=sum(lan_tot),fday=sum(fday))%>%
	mutate(perc_lan=round(100*lan_tot/sum(lan_tot),2))%>%
	ungroup()%>%
	arrange(desc(perc_lan))
kable(tab2)
```



Time series of landings in tons by area.

```{r explo2,fig.height=14,fig.width=8}
#landings by ICES division
tsarea<-laneff%>%
	group_by(area=sub("27.","",area),year)%>%
	summarise(lan_tot=sum(lan_tot,na.rm=T))
plt1<-ggplot(tsarea,aes(x=year,y=lan_tot))+
		geom_bar(stat="identity")+
		facet_grid(area~.,scale="free")+
		xlab("Year")+ylab("Landings (t)")+
		theme_bw()+
		ggtitle("French landings for red mullet")
print(plt1)
```

Landings by fishing gear and area 

```{r explo3,fig.height=14,fig.width=8}
#landings by gear
tsareamet<-laneff%>%
	group_by(area=sub("27.","",area),year,gear)%>%
	summarise(lan_tot=sum(lan_tot,na.rm=T))
plt1<-ggplot(tsareamet,aes(x=year,y=lan_tot,colour=gear,fill=gear))+
		geom_bar(stat="identity")+
		facet_grid(area~.,scale="free")+
		xlab("Year")+ylab("Landings (t)")+
		theme_bw()+
		ggtitle("French landings for red mullet")
print(plt1)
```

Idem with effort in fishing days

```{r explo4,fig.height=14,fig.width=8}
#landings by gear
tsareamet<-laneff%>%
	group_by(area=sub("27.","",area),year,gear)%>%
	summarise(fday=sum(fday))
plt1<-ggplot(tsareamet,aes(x=year,y=fday,colour=gear,fill=gear))+
		geom_bar(stat="identity")+
		facet_grid(area~.,scale="free")+
		xlab("Year")+ylab("Effort (fishing days)")+
		theme_bw()+
		ggtitle("French effort of vessels with red mullet landings")
print(plt1)
```

Map of lan and effort

```{r map1,fig.height=14,fig.width=14}
#load geographical object for mapping
#rect
load("../data/geo/rect.rdata")
#div
load("../data/geo/div.rdata")
#prepa geo data
rect<-rect%>%filter(ICESNAME%in%unique(laneff$rect))%>%
		transmute(rect=ICESNAME,div=F_DIVISION,label=ICESNAME,geometry)
rectxy<-st_coordinates(st_centroid(rect))
rect$x<-round(rectxy[,1],2)
rect$y<-round(rectxy[,2],2)
#div
div<-div%>%
	transmute(div=F_DIVISION,label=ETIQUETTE,geometry)%>%
	filter(div%in%unique(rect$div))
divxy<-st_coordinates(st_centroid(div))
div$x<-divxy[,1]
div$y<-divxy[,2]
#prep laneff for mapping
rlaneff<-laneff%>%group_by(year,rect)%>%
	summarise(lan_tot=sum(lan_tot),fday=sum(fday))%>%
	ungroup()%>%
	#remove empty rect
	filter(rect!="")%>%
	#add rect coordinates
	left_join(rect,by=c("rect"))
rx<-c(-13,7)#range(rlaneff$x)
ry<-c(43,55)#range(rlaneff$y)

pltlanyear<-ggplot()+
	geom_raster(data=rlaneff,aes(x=x,y=y,fill=lan_tot),stat="identity",alpha=1)+
	geom_sf(data=div,fill=NA,colour="black",lwd=.3,alpha=.5)+
	scale_fill_distiller(palette='Spectral',name="Landings (tons)")+
	facet_wrap(~year,drop=FALSE)+
	borders("world",fill="grey",colour=NA,alpha=.5)+
	coord_sf(xlim=rx,ylim=ry)+
	theme_bw()+
	ggtitle("French landings for red mullet")+
	xlab("Longitude")+ylab("Latitude")+
	theme(legend.position="bottom",
		      strip.text.y=element_text(size=6,angle=0))
print(pltlanyear)
plteffyear<-ggplot()+
	geom_raster(data=rlaneff,aes(x=x,y=y,fill=fday),stat="identity",alpha=1)+
	geom_sf(data=div,fill=NA,colour="black",lwd=.3,alpha=.5)+
	scale_fill_distiller(palette='Spectral',name="Landings (tons)")+
	facet_wrap(~year,drop=FALSE)+
	borders("world",fill="grey",colour=NA,alpha=.5)+
	coord_sf(xlim=rx,ylim=ry)+
	theme_bw()+
	ggtitle("Fishings days for french vessels with red mullet landings")+
	xlab("Longitude")+ylab("Latitude")+
	theme(legend.position="bottom",
		      strip.text.y=element_text(size=6,angle=0))
print(plteffyear)
```

Data extraction for maturity analysis

```{r dataextract,fig.height=14,fig.width=14}
matlaneff<-laneff%>%
		mutate(zone=NA)%>%
		mutate(zone=ifelse(area%in%c("27.8.a","27.8.b"),"8ab",zone))%>%
		mutate(zone=ifelse(area%in%c("27.7.d","27.4.c"),"7d4c",zone))%>%
	   group_by(year,zone)%>%
	   summarise(lan=sum(lan_tot),fday=sum(fday))%>%
	   ungroup()%>%
	   filter(!is.na(zone))%>%
	   mutate(cpue=lan/fday)
saveRDS(matlaneff,"../data_confidential/matlaneff.rds")
#format the data for easy ggploting
pipo<-pivot_longer(matlaneff,col=3:5)
 
#quick plot
ggplot(pipo,aes(x=year,y=value))+
	geom_point()+
	geom_path()+
	geom_smooth()+
	facet_wrap(name~zone,scale="free_y",ncol=2)

```
